\Title

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : b

We would like to highlight a few innovations in the contributions that are presented in this paper. First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif. Another novelty is that the model learns the weights and the delays simultaneously. For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning. In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}. Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs. Instead of relying on a careful description of the physical rules governing a task, e.g. the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified several limits to our model, which we will now detail. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations. % \note{pourrait parler du poster bernstein? de nadafian?}

Another restriction is that this model is purely feed-forward. Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field. However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}. For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}. Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, Notably to explain non-linear mechanisms~\citep{boutin_pooling_2022}. It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions. However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses. While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).  
% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}

>>>>

We would like to highlight a few innovations in the contributions that are presented in this paper. First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif. Another novelty is that the model learns the weights and the delays simultaneously. For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning. In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}. Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs. Instead of relying on a careful description of the physical rules governing a task, e.g. the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified a number of limitations of our model, which we will now discuss in detail. First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time. We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, and especially to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no effect on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes. This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme. In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}. Thus, it is possible to define a purely event-based framework. Such an architecture could provide promising computational speedups. % \note{pourrait parler du poster bernstein? de nadafian?}

A further limitation is that the model is purely feed-forward. Thus, the spikes generated by the postsynaptic neurons are based solely on information contained in the classical receptive field. However, it is well known that neurons in the same layer can interact with each other using lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}. For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}. Furthermore, neural information is modulated by feedback information, e.g. to distinguish a figure from its background~\citep{roelfsema_early_2016}. Feedback has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, especially to explain non-linear mechanisms~\citep{boutin_pooling_2022}. Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback). However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic theoretical model is able to incorporate them. While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
